% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode
% !BIB TS-program = bibtex
%
% Graph-signal-processing foundations for cycling-network expansion:
% spectral bounds, optimal siting and learning curves on the
% station-proximity graph of dock-based bike-share systems.
%
% Authors : R. Fossé & G. Pallares (CESI LINEACT, 2026-).
% Companion paper to the IMD-4 predictive paper (Fossé & Pallares 2026).
%
\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=2.5cm]{geometry}
\usepackage{microtype}
\usepackage{csquotes}

\usepackage{amssymb,amsmath,amstext,amsthm}

\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{float}
\usepackage{caption}
\captionsetup{font=small,labelfont=bf,labelsep=period}
\usepackage{enumitem}
\setlist{topsep=2pt,itemsep=2pt,parsep=0pt}
\usepackage{url}

\usepackage[authoryear,round]{natbib}
\bibliographystyle{elsarticle-harv}

\usepackage{xcolor}
\usepackage[colorlinks=true,
            linkcolor=black, citecolor=black, urlcolor=blue!50!black,
            breaklinks=true]{hyperref}

\usepackage{authblk}
\renewcommand\Authfont{\normalsize}
\renewcommand\Affilfont{\small\itshape}
\setlength{\affilsep}{0.6em}

\newcommand{\IMD}{\mathrm{IMD}}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newenvironment{highlights}{%
  \par\noindent\textbf{\large Highlights}\par\smallskip
  \begin{itemize}[leftmargin=1.4em,itemsep=0.15em,topsep=0.2em]}{%
  \end{itemize}\par\medskip}

% =========================================================================
\title{\Large\bfseries Graph-signal-processing foundations for\\
\large cycling-network expansion: spectral bounds, optimal siting and\\
empirical learning curves on dock-based bike-share systems}

\author[1]{Rohan Fossé\thanks{Corresponding author: \texttt{rfosse@cesi.fr}. ORCID: \href{https://orcid.org/0009-0002-2195-0198}{0009-0002-2195-0198}}}
\author[2]{Gaël Pallares\thanks{ORCID: \href{https://orcid.org/0009-0002-8680-604X}{0009-0002-8680-604X}}}
\affil[1]{CESI Engineering School, Montpellier, France}
\affil[2]{CESI LINEACT (EA 7527), Montpellier, France}

\date{Draft v0.1 — extracted 2026-05-21 from companion paper}

\begin{document}
\maketitle

% =========================================================================
\begin{abstract}
\noindent
\textbf{Working draft, v0.1 — extracted from the companion
predictive paper (Fossé \& Pallares, 2026) on 21 May 2026.}

Dock-based bike-share systems define a natural geometric object —
the \emph{station-proximity graph} — on which both supply-side
features (cycling infrastructure, transit multimodality, density)
and demand-side outcomes (hourly trip counts) live as graph
signals.  This companion paper develops the graph-signal-
processing (GSP) framework for that setting and uses it to derive
three classes of operational tools for cycling-network expansion:

\textbf{(I) Spectral diagnostics.}  We define the demand-signal
smoothness on the station-proximity Laplacian, derive an upper
bound on the in-sample $R^2$ recoverable from any linear
combination of $d$ supply-side features (Theorem 2), and
characterise the structural collapse of station-fixed-effect
models on held-out stations (Theorem 3).  Empirical measurements
on nine networks $\geq 400$ stations confirm the predicted
bounds.

\textbf{(II) Optimal siting via sampling theory.}  We cast new-
station siting as the $D$-optimal subset selection problem on
the station-proximity Laplacian eigenbasis and apply the
Nemhauser-Wolfe-Fisher submodular bound (Theorem 4) and the
Anis-Gadde-Ortega band-limited reconstruction guarantee
(Theorem 5).  Empirical comparison against MCLP and
$k$-median (Theorem~7's experimental table) shows that the
sampling-theoretic objective dominates rank-based metrics by
$\Delta \rho \in [+0.6, +0.9]$ at $K = 50$ on Boston and DC.

\textbf{(III) Learning curves and the saturation point.}  We
fit the empirical $\rho(N)$ saturation curve on Boston Bluebikes
($n_* = 21 \pm 3$, $\rho_\infty = 0.77$, Theorem 6) and use it
as a diagnostic to identify when a city has enough deployed
stations for IMD-4 transfer to be reliable.

We further benchmark the framework against a GCN baseline
\citep{Kipf2017}, a regularised-Laplacian smoother and Heat-
Kernel Signatures, and show that the GSP toolkit complements
rather than competes with the supply-side composite indicator
of the companion paper.

\noindent\textbf{Keywords:}
graph signal processing ; bike-sharing ; station siting ;
submodular optimisation ; spectral bounds ; sampling theory ;
learning curves.
\end{abstract}

\begin{highlights}
\item Spectral upper bound on the linear $R^2$ recoverable from $d$ supply-side features
\item Submodularity-based $1-1/e$ guarantee for $D$-optimal greedy station siting
\item Band-limited reconstruction bound (Anis-Gadde-Ortega) applied to bike-share demand
\item Empirical learning curve on Boston: $n_* = 21 \pm 3$ training stations to saturate
\item D-optimal siting beats MCLP / $k$-median by $\Delta \rho > +0.6$ at $K = 50$
\end{highlights}

% =========================================================================
\section{Introduction}
\label{sec:intro}

\textbf{[Draft placeholder — needs full rewrite as standalone.]}
This paper develops the graph-signal-processing (GSP) framework
that underpins the empirical results of the companion paper on
the IMD-4 cycling-environment indicator (Fossé \& Pallares,
2026, henceforth ``the predictive paper'').

The predictive paper documents two empirical findings that the
GSP framework makes structural :
\begin{itemize}[label=$\bullet$,leftmargin=*]
\item The IMD-4 features achieve Spearman $\rho \in [+0.49,
      +0.79]$ on leave-station-out predictions, while a station
      fixed-effect collapses to $\rho \in [-0.70, -0.32]$. The
      gap is a structural property of any signal-on-graph
      framework, not a property of the IMD-4 features
      specifically (Theorem 3).
\item The linear predictive ceiling on the 4-D IMD subspace is
      $R^2_{\mathrm{spectral}} \in [0.08, 0.32]$ across nine
      networks.  This is the spectral upper bound of Theorem 2,
      and it explains why non-linear gradient-boosted regressors
      do not recover more than $\sim 0.18$ out-of-sample R$^2$
      on station-mean demand from the IMD-4 axes alone (script
      d36 in the companion repository).
\end{itemize}

\textbf{Scope of this paper.}  Three operational use-cases
follow from the GSP framework, each addressed by a dedicated
tool : (i) \emph{diagnostic} — spectral concentration of demand,
smoothness and prediction ceilings on a candidate network ;
(ii) \emph{cold-start siting} — selecting $K$ new stations to
maximise the worst-case reconstruction error on the rest of the
candidate set ; (iii) \emph{infill} — predicting demand at unobserved
stations on a partly-deployed network via a regularised-Laplacian
smoother that fuses the IMD-4 prior with the realised demand at
deployed neighbours.  These three tools share the same graph
object and the same eigendecomposition ; this paper presents
them as a coherent GSP toolkit for cycling-network expansion.

\textbf{Relation to the companion paper.}  All empirical results
of this paper that involve trip-level data (Sections~\ref{sec:gsp-laplacian}, \ref{sec:gsp-siting}, \ref{sec:gsp-robustness})
re-use the data infrastructure of the companion paper
(Tier 1 / Tier 2 panels of nine LSO networks).  The IMD-4
indicator itself is defined and validated in the companion paper
and used here as a fixed supply-side input.

% =========================================================================
% §8 EXTRACTED VERBATIM FROM THE COMPANION PAPER (v2.3).
% =========================================================================

\input{_section8_extract}

% =========================================================================
\section{Discussion}
\label{sec:discussion}

\textbf{[Draft placeholder — needs writing as standalone.]}
Three takeaways for cycling-network expansion practice :

\begin{enumerate}[leftmargin=*]
\item The spectral upper bound of Theorem 2 sets a hard ceiling
      on what \emph{any} linear combination of $d$ supply-side
      features can recover from demand on a smooth graph.  A
      city operator inspecting a candidate feature set can
      compute $R^2_{\mathrm{spectral}}$ ahead of deployment and
      decide whether the feature set is sufficient.
\item The D-optimal siting of Section~\ref{sec:gsp-siting}
      provides a $(1-1/e)$-approximation guarantee under
      submodularity (Theorem 4) and operationally beats MCLP /
      $k$-median by a wide margin on the bike-share use-case,
      because it optimises against demand prediction error
      rather than against a flat coverage metric.
\item The empirical learning curve $\rho(N) = \rho_\infty
      (1 - \mathrm{e}^{-N/n_*})$ on Boston gives a sample-size
      diagnostic : with $n_* \approx 21$, an operator can
      identify when an existing partial deployment carries
      enough information for IMD-4 LSO transfer to be reliable.
\end{enumerate}

% =========================================================================
\section{Limits}
\label{sec:limits}

\textbf{[Draft placeholder.]}

\begin{itemize}[label=$\bullet$,leftmargin=*]
\item \textbf{Theorems 1--5 are applications of standard GSP /
      OR results to the bike-share setting} rather than new
      mathematical theorems.  T1 is the Dirichlet-energy /
      uniform-permutation bound, T2 is the orthogonal-projection
      $R^2$, T3 is structural, T4 is Nemhauser-Wolfe-Fisher
      (1978), T5 is Anis-Gadde-Ortega (2016).  The empirical
      novelty is the systematic application to the bike-share
      use-case ; the theoretical content is mostly synthesis.
      Theorem 6 (learning curve) is fitted on Boston only and
      requires replication on more cities for a
      population-level claim.
\item \textbf{$k$-NN graph parameter ($k = 6$) and Gaussian
      kernel bandwidth ($\sigma = 300$\,m) not yet sensitivity-
      tested.}  Both are hard-coded to match the IMD-4 buffer
      length scale.
\item \textbf{MCLP comparison} reported on Boston + DC only at a
      single coverage radius ($500$\,m).
\item \textbf{GCN baseline} is a $2$-layer Kipf-Welling model
      with default hyperparameters ; ablations on depth, dropout
      and edge-weight schemes are not reported.
\end{itemize}

% =========================================================================
\section{Code and data}

All scripts referenced in this paper are in the
\texttt{experiments/} subdirectory of the
\texttt{paper\_gsp\_companion/} folder of the
\href{https://github.com/rohanfosse/imd-national-catalogue}{imd-national-catalogue}
repository :

\begin{itemize}[leftmargin=*,itemsep=0pt]
\item \texttt{d22\_lso\_robustness.py} — robustness of LSO to spatial-block folds
\item \texttt{d23\_synthetic\_graph\_city.py} — synthetic-city verification of the spectral bound
\item \texttt{d24\_gsp\_real\_cities.py} — empirical GSP measurements on the 9 LSO networks
\item \texttt{d25\_graph\_laplacian\_lso.py} — regularised-Laplacian smoother
\item \texttt{d26\_optimal\_siting.py} — D-optimal greedy siting
\item \texttt{d27\_ga\_imd.py} — Graph-augmented IMD-4 hybrid
\item \texttt{d28\_spectral\_bottleneck.py} — $R^2_{\mathrm{spectral}}$ computation
\item \texttt{d29\_kernel\_zoo.py} — kernel-family sensitivity
\item \texttt{d30\_gcn\_baseline.py} — GCN baseline
\item \texttt{d31\_heat\_kernel\_signatures.py} — HKS features
\item \texttt{d32\_learning\_curve.py} — saturation curve on Boston
\item \texttt{d33\_mclp\_comparison.py} — MCLP / $k$-median baselines
\item \texttt{d36\_nonlinear\_R2\_spectral.py} — non-linear ceiling
\item \texttt{d40\_paris\_spectral\_3axes.py} — Paris spectral fix
\end{itemize}

\bibliography{references/references_gsp}

\end{document}
